---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The data I am analyzing is the books checked out more than ten times from 2017-2023. One thing I want to find out is which medium or UsageClass is more popular and if the ratio changes from year to year. I would also like to do something similar for the MaterialType of the publication like ebook, magazine, movie, music, etc. The last thing I would like to analyze is what books are the most popular and how much does that list change from year to year.

### Summary Information

```{r, echo = F, message=FALSE, warning=FALSE, code = readLines("summary.R")}
```
There are `r num_digital_checkouts` digital checkouts within our dataset and `r num_physical_checkouts` physical checkouts in the dataset which split the `r total_checkouts` total checkouts evenly. Also I found that there are `r publication_types` types of publications like Ebooks, Magazines, Books, videos, etc. Another interesting piece of my data that I found is that the most viewed book overall is `r most_viewed`

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

``````{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart-1.R")}
```
This chart shows the total number of checkouts per year based on whether the checkout was digital or physical. I think it is kind of interesting because the trend does shift quite drastically and at first I didn't understand why. After some thought, I realized the years when digital overtook Physical were the same years that the Pandemic was at its peak. 2023 has only just begun, but you can still see that the trend from the Pandemic has continued and there are still more Digital checkouts than Physical.

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```


